An Algebraic Theory of File Synchronization

### 2015-12-25 2015/def-commands.tex
More improvements

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work
demonstrated (\cite{BZ})
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depend on the types of values in the filesystem.

>>>>

Let us consider what kind of information is usually encoded in filesystem operations.
A minimal set of commands, based on the most frequent tools implemented by filesystems,
may be $create(n,\valv)$, which creates a file or directory ($\valv$) at the empty node $n$;
$edit(n,\valv)$, which replaces the earlier value at $n$ with $\valv$;
and $remove(n)$, which deletes the value at $n$---where 
$n\in\setn$ and $\valv\in\setv$ (but $\valv\neq\empt$).
Regarding their output, that is, the state of the filesystem at $n$
after applying the command,
we know that after $create$ or $edit$, $\FS(n)\neq\empt$, whereas after $remove$,
$\FS(n)$ will be $\empt$. 
However, from \cite{NREC} and \cite{CBNR} we know that a useful set of axioms
will in some cases need to distinguish between, for example,
$edit$s that result in directories ($edit(n,\vald)$) and
ones that result in files ($edit(n,\valf)$), and treat them as separate commands,
as their behaviors are quite different when combined with other commands.
Indeed, Bill Zissimopoulos' work
demonstrated (\cite{BZ})
that extending this distinction to more commands ultimately simplifies
the definition of conflicting commands, as our model will then able to predict the behavior of commands
more precisely.
In other words, encoding the type of the output ($\cchard$, $\ccharf$ or $\ccharb$) in the commands is definitely useful.
At the same time, there is never any need to consider the
exact output \emph{value} of a command,
as the success or failure of filesystem commands only depends on the types of values in the filesystem.

### 2015-12-25 2015/intro.tex
More improvements

In this paper we build on the work done 
on command-based algebraic file synchronization
in \cite{NREC} and add to the theoretical understanding
of synchronization by providing rigorous proofs that the algorithms
we describe work as intended
and offer a solution in our framework that cannot be improved on.

A central problem of command-based synchronizers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems without causing an error 
(for example, a directory needs to be created before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.

>>>>

In this paper we build on the work done 
on an algebraic approach to synchronization
in \cite{NREC} and add to the theoretical understanding
of synchronization by providing rigorous proofs that the algorithms
we describe work as intended
and offer a solution in our framework that cannot be improved on.

A central problem of command-based synchronizers 
during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied
to the filesystems without causing an error 
(for example, a directory needs to be created before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands (updates),
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.
Other research also shows the essential nature of ordering:
in IceCube multiple orders are tested to find an acceptable one \cite{KRSD},
or in Bayou reconciling updates happens by redoing them in an acceptable order \cite{TTPDSH}.
